Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-16 ICML-submission/introduction.tex
Fix introduction

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}, formally
stated in
Definitions~\ref{definition:weak-linear-separability}~and~\ref{definition:strong-linear-separability}.
The standard notion of multiclass linear separability corresponds to the weak
linear separability. For binary classification, the weak and strong separability
are identical. However, they differ for multiclass classification with $K \ge 3$
classes. For multiclass classification with $K$ classes, the standard notion of
linear separability means that the examples from each class lie in an
intersection of $K-1$ halfspaces and the examples outside of the class lie in
the complement of the intersection of the halfspaces. Strong linear separability
means that examples from each class are separated from the remaining examples by
a \emph{single} hyperplane.

In the full-information feedback setting, it is well known that if all examples
have norm at most $R$ and are weakly linearly separable with a margin $\gamma$,
then the \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes. The guarantees provided by \textsc{Multiclass
Perceptron} is very satisfying, since it achieves an information-theoretically
optimal mistake upper bound, while having low time and memory complexities.
\footnote{These results are standard in the literature. For completeness, we
present Theorems~\ref{theorem:mutliclass-perceptron-mistake-upper-bound}
and~\ref{theorem:online-multiclass-classification-mistake-lower-bound}, along
with their proofs in Appendix~\ref{section:multiclass-perceptron-proofs}.}

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work. \footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect. Their Lemma 4 cannot be directly applied to the bandit
setting, as the algorithm only makes updates on a subset of the $w_i$'s.} We
design a simple and efficient algorithm, namely
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
that has an expected number of mistakes at most $(K-1) \lfloor 4(R/\gamma)^2
\rfloor$. The algorithm can be viewed as running $K$ copies of the
\textsc{Binary Perceptron} algorithm, one copy for each class. Intuitively
speaking, the $O(K)$ multiplicative factor is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We also
prove that any (possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the case when examples are weakly linearly separable, we construct a feature
mapping $\phi$ that makes them \emph{strongly} linearly separable. We then use
the kernelized version of the algorithm, namely
Algorithm~\ref{algorithm:kernelized}, for the strongly separable case; for more
details on kernel methods see e.g.~\citet{Scholkopf-Smola-2002} and
\citet{Shawe-Taylor-Cristianini-2004}. The kernel $k(x,x')$ corresponding to the
feature mapping $\phi$ has a simple explicit formula and can be computed in
$O(d)$ time.

The number of mistakes of the kernelized algorithm depends on the margin in the
corresponding feature space. We analyze how the margin parameter of weak
separability in the original space $\R^d$ gets transformed into a margin
parameter of strong separability in the new feature space. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \citet{Klivans-Servedio-2008}. As a side result, we
improve on the results of \citet{Klivans-Servedio-2008} by removing the
dependency on the original dimension $d$.

>>>>

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}, formally
stated in
Definitions~\ref{definition:weak-linear-separability}~and~\ref{definition:strong-linear-separability}.
The standard notion of multiclass linear separability corresponds to the weak
linear separability. For binary classification, the weak and strong separability
are identical. However, they differ for multiclass classification with $K \ge 3$
classes. For multiclass classification with $K$ classes, weak linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane.

In the full-information feedback setting, it is well known that if all examples
have norm at most $R$ and are weakly linearly separable with a margin $\gamma$,
then the \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes. The guarantees provided by \textsc{Multiclass
Perceptron} is very satisfying, since it achieves an information-theoretically
optimal mistake upper bound, while having low time and memory
complexities.~\footnote{These results are standard in the literature. For
completeness, we present these results along with their proofs in
Appendix~\ref{section:multiclass-perceptron-proofs} in the supplementary
material.}

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect. Their Lemma 4 cannot be directly applied to the bandit
setting, as the algorithm only makes updates on a subset of the $w_i$'s.} We
design a simple and efficient algorithm, namely
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
that has an expected number of mistakes at most $(K-1) \lfloor 4(R/\gamma)^2
\rfloor$. The memory complexity and per-round time complexity are both $O(dK)$.
The algorithm can be viewed as running $K$ copies of the \textsc{Binary
Perceptron} algorithm, one copy for each class. Intuitively speaking, the $O(K)$
multiplicative factor in the mistake bound is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We prove
that any (possibly randomized) algorithm must make $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the case when examples are weakly linearly separable, we construct a
non-linear feature mapping $\phi$ and associated positive definite kernel
$k(x,x')$ that makes the examples \emph{strongly} linearly separable. We then
use the kernelized version of the algorithm for the strongly separable case,
Algorithm~\ref{algorithm:kernelized}. (For details on kernel methods see
e.g.~\citet{Scholkopf-Smola-2002} or \citet{Shawe-Taylor-Cristianini-2004}.)
The kernel $k(x,x')$ corresponding to the feature mapping $\phi$ has a simple
explicit formula and can be computed in $O(d)$ time.

The number of mistakes of the kernelized algorithm depends on the margin in the
corresponding feature space. We analyze how the mapping $\phi$ transforms the
margin parameter of weak separability in the original space $\R^d$ into a margin
parameter of strong separability in the new feature space. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \citet{Klivans-Servedio-2008}. As a side result, we
improve on the results of \citet{Klivans-Servedio-2008} by removing the
dependency on the original dimension $d$.
